Conversations in time: interactive visualisation to explore structured temporal data

### 2020-09-07 tsibbletalk.Rmd
done with slicing and dicing

This shiny module, decoupled to `tsibbleDiceUI()` and `tsibbleDiceServer()`, presents a clean interface and forms a resusable piece in a shiny application. Like all shiny modules, the first argument in both functions requires a user-supplied session id that must be unique. The UI function `tsibbleDiceUI()` simply shows a slider that animates or controls the number of periods to be diced. The workhorse is certainly the server function `tsibbleDiceServer()`, encapsulating the core algorithm that transforms data and sends messages to update the plot accordingly. It expects a `ggplot` or `plotly` object, where one can plot data using either lines or other graphical elements (such as boxplots). As the function name suggests, a (shared) tsibble is needed to start the engine. The `period` argument is to specify the minimum number of seasonal blocks, for example data shifted by "1 day", "2 days", or "1 week", etc. The following code chunk generates Figure \ref{fig:wrap-ped}.

@@

The initial state is

(Data flows) Transformed data generally requires redrawing the plot, and worsen the performance of shiny. The underlying tsibble data is called back and transformed in R. Using the `plotly.js` react method, only transformed data is sent to the server side, while keeping the rest configuration unchanged (e.g. layout and graphical elements). It is performant, and users will not experience notable delay in response to the change in the slider input. Dissect time index, and propagate transformed data to shiny server.

@@

# References {-}

>>>>

This shiny module, decoupled to `tsibbleDiceUI()` and `tsibbleDiceServer()`, presents a clean interface and forms a resusable piece in a shiny application. Like all shiny modules, the first argument in both functions requires a user-supplied session id that must be unique. The UI function `tsibbleDiceUI()` simply shows a slider that animates or controls the number of periods to be diced. The workhorse is certainly the server function `tsibbleDiceServer()`, encapsulating the core algorithm that transforms data and sends messages to update the plot accordingly. It expects a `ggplot` or `plotly` object, where one can plot data using either lines or other graphical elements (such as boxplots). As the function name suggests, a (shared) tsibble is needed to start the engine, and thereby the time `index` can be retrieved for dissection. The `period` argument is to specify the minimum number of seasonal blocks as desired, for example data shifted by "1 day", "2 days", or "1 week", etc. The following code chunk generates Figure \ref{fig:wrap-ped}.

@@

Upon running the shiny application, Figure \ref{fig:wrap-ped-1} corresponds to the initial state, with the slider incremented by 1-day unit. The "play" button near the end of slider begins animating the slicing and dicing process by walking through all 24 hours by 153 days. Alternatively, users can drag the handler to poke around certain slices themselves.

In response to the slider input, the plot will be updated and loaded with newly transformed data. Keeping the application as performant as possible is the top priority. Without completely redrawing the plot, the `plotly.js` react method is invoked internally. The underlying tsibble data is being called back and processed in R. Only transformed data gets fed back to the shiny server, along with reseting the x-axis ranges and breaks. The rest plot configurations, such as marks, y-axes, and layouts, are properly cached.

The new shiny module exploits the temporal aspect for a tsibble object. It allows users to slide through relative periods to explore seasonal insights, with slick user experience.

@@

# References

### 2020-09-07 tsibbletalk.tex
done with slicing and dicing

This shiny module, decoupled to \texttt{tsibbleDiceUI()} and
\texttt{tsibbleDiceServer()}, presents a clean interface and forms a
resusable piece in a shiny application. Like all shiny modules, the
first argument in both functions requires a user-supplied session id
that must be unique. The UI function \texttt{tsibbleDiceUI()} simply
shows a slider that animates or controls the number of periods to be
diced. The workhorse is certainly the server function
\texttt{tsibbleDiceServer()}, encapsulating the core algorithm that
transforms data and sends messages to update the plot accordingly. It
expects a \texttt{ggplot} or \texttt{plotly} object, where one can plot
data using either lines or other graphical elements (such as boxplots).
As the function name suggests, a (shared) tsibble is needed to start the
engine. The \texttt{period} argument is to specify the minimum number of
seasonal blocks, for example data shifted by ``1 day'', ``2 days'', or
``1 week'', etc. The following code chunk generates Figure
\ref{fig:wrap-ped}.

@@

The initial state is

(Data flows) Transformed data generally requires redrawing the plot, and
worsen the performance of shiny. The underlying tsibble data is called
back and transformed in R. Using the \texttt{plotly.js} react method,
only transformed data is sent to the server side, while keeping the rest
configuration unchanged (e.g.~layout and graphical elements). It is
performant, and users will not experience notable delay in response to
the change in the slider input. Dissect time index, and propagate
transformed data to shiny server.

>>>>

This shiny module, decoupled to \texttt{tsibbleDiceUI()} and
\texttt{tsibbleDiceServer()}, presents a clean interface and forms a
resusable piece in a shiny application. Like all shiny modules, the
first argument in both functions requires a user-supplied session id
that must be unique. The UI function \texttt{tsibbleDiceUI()} simply
shows a slider that animates or controls the number of periods to be
diced. The workhorse is certainly the server function
\texttt{tsibbleDiceServer()}, encapsulating the core algorithm that
transforms data and sends messages to update the plot accordingly. It
expects a \texttt{ggplot} or \texttt{plotly} object, where one can plot
data using either lines or other graphical elements (such as boxplots).
As the function name suggests, a (shared) tsibble is needed to start the
engine, and thereby the time \texttt{index} can be retrieved for
dissection. The \texttt{period} argument is to specify the minimum
number of seasonal blocks as desired, for example data shifted by ``1
day'', ``2 days'', or ``1 week'', etc. The following code chunk
generates Figure \ref{fig:wrap-ped}.

@@

Upon running the shiny application, Figure \ref{fig:wrap-ped-1}
corresponds to the initial state, with the slider incremented by 1-day
unit. The ``play'' button near the end of slider begins animating the
slicing and dicing process by walking through all 24 hours by 153 days.
Alternatively, users can drag the handler to poke around certain slices
themselves.

In response to the slider input, the plot will be updated and loaded
with newly transformed data. Keeping the application as performant as
possible is the top priority. Without completely redrawing the plot, the
\texttt{plotly.js} react method is invoked internally. The underlying
tsibble data is being called back and processed in R. Only transformed
data gets fed back to the shiny server, along with reseting the x-axis
ranges and breaks. The rest plot configurations, such as marks, y-axes,
and layouts, are properly cached.

The new shiny module exploits the temporal aspect for a tsibble object.
It allows users to slide through relative periods to explore seasonal
insights, with slick user experience.
